---
title: "Case Template"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    embed-resources: true
execute:
  include: true
  eval: true    
  warning: false
  message: false

---
We start by loading the necessary libraries and the datasets.


# Q1

If you repeat the question, do not make it part of the heading!

```{r}
# code annotations go within the chunks

2 + 2 # example of addition
```

Comments on your results go here:

- Bullets help organize your material
- For bullets to work there needs to be a space after the preceding line
- Emphasis can be added with *italics* and **bolding**

If you want to highlight your answer (without adding it to the ToC) simply bold it:

**Answer**:

- point 1
- point 2

The code chunks included here are R chunks but you could use python just as easily by changing the chunk instructions:  `{python}`.

# Q2

```{r}

```

# Q3

```{r}

```

# Q4

```{r}

```

# Q5

```{r}

```

# Q6

```{r}

```

# Q7

```{r}

```

